---
title: "BST-262: Assignment 1"
author: "Daniel Briggs"
date: "Due Friday, November 30 2018, 5pm"
output: pdf_document
---




```{r}
library(tidyverse)
library(fiftystater)
library(knitr)
```



# Exercise 1 - Julia (language: Julia)

Here for any dictionary, we will take the maximum value's (assuming numeric values) index, and return the key which matches that value.

```
function getmax(d)
    maxind = argmax(collect(values(d))) #identify the index of the maximum value
    return collect(keys(d))[maxind] #indices the key by the indexed max
end
```

# Exercise 2 - Data parallelism vs task parallism (language: R)

## Question 1

Add comments

```{r}
# Add your R code
library(tm)
library(doParallel)

#patterns for mapping
patterns <- c('ARTHROTEC', 'LIPITOR', 'VOLTAREN')

#initializing vectors to hold the data
proportions <- sums <- rep(NA,length(patterns))

#extra index used for assignment
k <- 1
for(i in patterns){
  
  #initalize a corpus based on the pattern
  cadec <- VCorpus(DirSource(directory = "data/cadec/text/", encoding = "UTF-8",pattern = i))
  
  #how many of each type?
  sums[k] <- length(names(cadec))
  
  #getting the text vectors
  cadec_content <- sapply(content(cadec), content)

  #creating the cluster
  cl <- makeCluster(detectCores())
  
  #assignment of tasks among the cluster
  cadec_part <- clusterSplit(cl,seq_along(cadec))
  cadec_partitioned <- lapply(cadec_part, function(p) cadec_content[p])
  
  #parallel saply to find pattern
  l <- parSapply(cl, cadec_partitioned, grepl, pattern = "pain")

  proportions[k] <- sum(parSapply(cl, l, sum))/sums[k]

  stopCluster(cl)
  
  k <- k+1
}


data.frame(drugs= patterns, total = sums, proportion = proportions)
```

## Question 2

Add comments

```{r}
parallel_pattern_sums <- function(pattern, vCorp){ #wrapper function
  corpus_content <- sapply(content(vCorp),content) #corpus content extraction
  
  cl <- makeCluster(detectCores()) #build cluster with cores
  
  corpus_part <- clusterSplit(cl, seq_along(corpus_content)) #partition tasks to cores

  corpus_partioned <- lapply(corpus_part, function(p) corpus_content[p]) #grabs content from specified divided corpus
  
  l <- parSapply(cl,corpus_partioned,grepl, pattern = pattern) #parallel sapply
  sum <- sum(parSapply(cl,l,sum)) #get the sum
  
  stopCluster(cl) #kills cluster
  return(sum)
}

#execute code
vc <- VCorpus(DirSource(directory = "data/cadec/text/", encoding = "UTF-8"))
patterns2 <- c('pain', 'bleeding', 'headache', 'nausea')

#get the sums
sums <- base::sapply(patterns2, parallel_pattern_sums,vCorp = vc)
print(sums)

#sum the sums
sum(sums)
```



# Exercise 3 - Grouped data operations and maps (language: R or Python)

## Question 1

Add comments

```{r}
# read in the csv
inpat <- read_csv("data/medicare/Medicare_Provider_Charge_Inpatient_DRGALL_FY2015.csv")

# determine the total cost by state
state_inpat <- inpat %>% mutate(`Total Cost` = `Total Discharges` * `Average Total Payments`) %>% group_by(`Provider State`) %>% summarize(`In Cost` = sum(`Total Cost`))

#print the resulting data frame
state_inpat
```


## Question 2

Add comments

```{r}
#read in the csv
outpat <- read_csv('data/medicare/Medicare_Charge_Outpatient_APC28_CY2015_Provider.csv')

#makes the names consistent across data frames
names(outpat) <- gsub(pattern = "_",replacement = " ",x = names(outpat))

#create the new data frame grouped by state and determining the total cost of services in each state
state_outpat <- outpat %>% mutate(`Total Cost` = `Outpatient Services` * `Average Total Payments`) %>% group_by(`Provider State`) %>% summarize(`Out Cost` = sum(`Total Cost`))

#print out the data set
state_outpat
```


## Question 3

Add comments

```{r}
#combine the data sets
Merged_States <- state_inpat %>% inner_join(state_outpat) %>% mutate(`Total Cost` = `In Cost` + `Out Cost`)

#manual processing of states
states <- tolower(state.name[match(state.abb, Merged_States$`Provider State`)])
states[which(is.na(states))] <- "district of columbia"
Merged_States$`Provider State` <- states

#read in map data
data("fifty_states")

#create the map
ggplot(Merged_States, aes(map_id = `Provider State`)) +  #give the data ggplot
  geom_map(aes(fill = log(`Total Cost`)), map = fifty_states) + #add the map layer
  expand_limits(x = fifty_states$long, y = fifty_states$lat) + #make the limits according to lat and longitude
  coord_map() + #project the map onto a 2D plane
  scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL) + #remove X/Y axis ticks ecause they are ugly
  labs(x = "", y = "") + theme(legend.position = "left", 
        panel.background = element_blank()) 

```


